---
title: "04_RMarkdown"
author: "Sachida Pokhrel"
date: "2025-02-26"
output: 
  html_document:
    toc: true
    toc_float: true
  md_document:
    variant: gfm
  word_document:
  pdf_document:
    
---

This is an example of R code.
```{r R code example, include=FALSE, echo=TRUE}
#include FALSE donot show anything regardless of echo. Echo decides if we can show code or not in the knitted document.If we use include TRUE and echo FALSE, it will not show code but show output.Default is true for both that means shows code and the output
#this is R code
summary(mtcars)
```


This is how we can include figures
```{r Include figures}
library(ggplot2)
data("mtcars")
ggplot(mtcars, aes(x=wt, y=mpg))+
  geom_point()
```

R markdown formatting options

# First level header

## Second level header
#### sub header for second level header
### Third level header

*this text is italics* #single asteric for italics beginning and at the end of line
**This text is bold** #double asteric to make it bold

- one item
- another item
  - sub item #tab once to make a subitem

## Links

[Link to my github](https://github.com/SachidaPokhrel/Class_Reproducibility/)
#[] provide the name to the link
## Images
![Shrek ggplot] (shrek.jpg) # ! specifies that we want to insert image.

## Formatted tables
```{r}
#install.packages("kable")
library(knitr)
kable(mtcars, digits = 3, format = "markdown") # kable brings table and digits 3 gives 3 digits after decimal
```

