---
title: "07_InClass_Assignment_LinearModels"
author: "Sachida Pokhrel"
date: "2025-04-02"
output: 
  md_document:
    variant: gfm
  html_document:
    toc: TRUE
    toc_float: TRUE
---

### Question 1
Read in the data called “PlantEmergence.csv” using a relative file path and load the following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting  and Rep into factors using the function as.factor
```{r}
STAND <- read.csv("07_LinearModels/PlantEmergence.csv")
library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
library(multcompView)

STAND$Treatment <- as.factor(STAND$Treatment)
STAND$DaysAfterPlanting <- as.factor(STAND$DaysAfterPlanting)
STAND$Rep <- as.factor(STAND$Rep)
```


### Question 2
Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results. 
```{r}
results <- lm(Emergence~Treatment*DaysAfterPlanting, data = STAND)
summary(results)
anova(results)
```
### Question 3
Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.

```{r}
results1 <- lm(Emergence~Treatment+DaysAfterPlanting, data = STAND)
summary(results1)
confint(results1)
anova(results1)
str(STAND)
```
Since the interaction is not significant when the interaction was included in the model, we donot need to include interaction (p-value =1). 

The intercept is the average emergence of Treatment1 which was 182.163(+/- 6.7753, 95% CI) . We found that the Treatment2 had 134.531 lesser emergence than Treatment1.(p-value = < 2e-16)

### Question 4
Calculate the least square means for Treatment using the emmeans package and perform a Tukey separation with the compact letter display using the cld function. Interpret the results.
```{r}
lsmeans <- emmeans(results1, ~Treatment)
results_lsmeans <- cld(lsmeans, alpha = 0.05, details = TRUE)
results_lsmeans
```

Different letters designate the significantly different emergence between the treatment. If some treatment share same letters, they are not statistically different. 
Treatment2 has the least emergence among the compared treatments which was significantly different from all the rest of the treatments. However, the treatment 1, 4, 8 and 9 that shows moderate emergence donot have statisticially significant difference in the emergence from Treatment 3,5,6 or  7 since they all share common letters.
The highest emergence was present in Treatment 5 which was statistically different from  Treatment 2 and 7 but not with the rest of the treatment.

### Question 5
The provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters. 
```{r}
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]

  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor))) # estimate lsmeans 
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE, Letters = letters) # contrast with Tukey adjustment by default.
  
  # Extracting the letters for the bars
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1], 
                                 str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")
  
  # for plotting with letters from significance test
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) + 
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") + 
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  
  return(plot)
}

plot_cldbars_onefactor(results1, "Treatment")

```

Here different letter signifies the statistical difference in the Treatment. Here Treatment2 that has the lowest emergence having significant letter "c" is different from the rest of the treatments. However, the treament 1, 4, 8 and 9  donot have statistically significant difference in the emergence from Treatment 3,5,6 or  7 since they all share common letters. For example; The treatment 4 has "ab" as a group which is not significantly different from Treatment 3 that has "a" or Treatment 7 that has "b" since they share common letters "a" or "b". However, if we are comparing the Treatment 3 and Treatment 7, they are different since they have different significant letters.

### Question 6
Generate the gfm .md file along with a .html, .docx, or .pdf. Commit, and push the .md file to github and turn in the .html, .docx, or .pdf to Canvas. Provide me a link here to your github.

**07_LinearModels** has file name **07_InClass_Assignment_LinearModels**

[Link to my Github](https://github.com/SachidaPokhrel/Class_Reproducibility)
