---
title: "05_InClass_Assignment_DataWrangling"
author: "Sachida Pokhrel"
date: "2025-03-20"
output: 
  html_document:
  md_document:
    variant: gfm
---
### Question 1. 
Read Data
```{r Read Data and load library}
#load data
Metadata <- read.csv("05_DataWrangling/Metadata.csv", na.strings = "na")
DiversityData <- read.csv("05_DataWrangling/DiversityData.csv", na.strings = "na")
#load library
library(tidyverse)
```
### Question 2. 
Combine two dataset with one common column "Code"
```{r Using left_join}
alpha <- left_join(Metadata, DiversityData, by = "Code")
```

### Question 3. 
Calculate Pielou's evenness index: Shannon/log(richness)
```{r using mutate}
alpha_even = mutate(alpha, even = shannon/log(richness))
```

### Question 4. 
Use mutate function to make a new dataset with summarise function and piping in tidyverse
```{r Grouping and summarising}
alpha_average <- alpha_even %>% 
  group_by(Crop, Time_Point) %>% 
  summarise(mean.even = mean(even),
            n = n(),
            sd.dev=sd(even)) %>% 
  mutate(std.err = sd.dev/sqrt(n))
alpha_average
```

### Question 5. 
Calculate the difference between the soybean column, the soil column, and the difference between the cotton column and the soil column
```{r Selecting certain column and making dataset wider and calculating difference}
alpha_average2 <- alpha_average %>% 
  select(Time_Point, Crop, mean.even) %>% 
  pivot_wider(names_from = Crop, values_from = mean.even) %>% 
  mutate(diff.cotton.even = Soil - Cotton) %>% 
  mutate(diff.soybean.even = Soil - Soybean)
```

### Question 6. 
Integrating the data into plot after converting the data into the longer version for the differences
```{r Making datset longer format and integrating into the plot}
#load library
library(ggplot2)
alpha_average2 %>% 
  select(Time_Point, diff.cotton.even, diff.soybean.even) %>% 
  pivot_longer(c(diff.cotton.even, diff.soybean.even), names_to = "diff") %>% 
  ggplot(aes(x=Time_Point, y=value, color = diff))+
  geom_line(stat = "identity")+
  xlab("Time (hrs)")+
  ylab("Difference from soil in Pielou's evenness")+
  theme_classic()
```

### Question 7.
#### Push to github
##### The folder is **05_DataWrangling** and the assignment has the file name **InClass_Assignment_DataWrangling.md**

[Link to my GitHub](https://github.com/SachidaPokhrel/Class_Reproducibility/tree/main)


